---
title: "baseline-mutation-rate"
output: html_document
---

```{r, echo=T, warning=F, message=F}
library(tidyverse)
library(ggplot2)
```

 Consider 3-mer and 7-mer mutation rate, see how they differ from each other. 
 
 
## 3-mer base rate 



```{r, echo=F}
three_mer_rate=as_tibble(read.table("C:\\Shengtong\\Research\\mutation-rate\\3_mer_base_from_Exac\\forweb_cleaned_exac_r03_march16_z_data_pLI.txt", header=T))
```

```{r, echo=T}
three_mer_rate
colnames(three_mer_rate)
```

Data (shouldn't be 3-mer rate) is downloaded from [here](ftp://ftp.broadinstitute.org/pub/ExAC_release/release1/manuscript_data/forweb_cleaned_exac_r03_march16_z_data_pLI.txt.gz)




## 3-mer vs 7-mer 

```{r, echo=F}
mer3_vs_mer7=as_tibble(read.table("C:\\Shengtong\\Research\\mutation-rate\\mutation-rate-modeling\\data\\mutation.rate.3mer.vs.7mer.random.10000row", header=F))
# originally from /project2/xinhe/Shengtong/mutation_rate/mutation.rate/baseline.mutation/b# aseline.comparison.between.3mer.and.7mer 

ggplot(mer3_vs_mer7, aes(x = V1, y = V2)) +
    geom_point()+
  xlab("3-mer mutation rate")+
  ylab("7-mer mutation rate")+
  geom_abline(intercept = 0, slope=1, col="red")+
  ggtitle("chr10 with alt_A")+
  theme(plot.title = element_text(hjust = 0.5)) # center the title 
```


* randomly choose 10,000 positions on chr10 with alternative allele `A`, compare 3-mer mutation rate and 7-mer rate. They appear to be identical.   

