---
title: "De novo mutations of ASD"
author: "Shengtong  Han"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

## Samples and reference windows


### Reference window

Example_windows_with_div_score.bed: each row is a 50bp window with a few features, such as mutation rate, coding, promoter, GC content, and divergence score; only cover genes and 10-kb regions within TSSs. 


### Case Samples 



index    | Samples | \# mutations| \# subjects
----|-------|-------|---
1 | Jiang_cases_DNM.bed | 2,091| 32
2 | Michaelson_cases_DNM.bed |581 | 10
3 | Yuen_NM2015_cases_DNM.bed | 9,381| 162
4 | Wu_cases_DNM.bed | 1,915 | 32
5 | Kong_cases_DNM.bed | 4,930 | 78

Table: Five WGS  DNM case sample information



### Control sample 

index    | Samples | \# mutations| \# subjects
----|-------|-------|---
1 | Wong_NC_2016_693_control_SNV.bed | 27,092|  693
Table:WGS  DNM control sample information 


### use reference window of 50 bp


```{r, echo=F}
ref.50bp.win=read.table("D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\Wong_NC_2016_693_control_SNV.Example_windows_with_div_score.bed", header=F)
unique.win.start=unique(ref.50bp.win[,5])
for ( i in 1:length(unique.win.start))
{
  i=1
  mutation.data=ref.50bp.win[ref.50bp.win[,5]==unique.win.start[i],]
  count=nrow(mutation.data)
  x1=mutation.data[,11]; x2=mutation.data[,12]; x3=mutation.data[,13]
  fit <- glm(count ~ x1+x2+x3, family=poisson())
}
```
<!-- Add your analysis here -->


### 1kb window 

```{r, echo=F}
ref.1kbp.win=read.table("D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\hg19.genome.1000bp.window.Wong_NC_2016_693_control_SNV.Example_windows_with_div_score.bed", header=F)
unique.win.start=unique(ref.1kbp.win[,2])
for ( i in 1:length(unique.win.start))
{
  i=2
  mutation.data=ref.1kbp.win[ref.1kbp.win[,2]==unique.win.start[i],]
  count=rep(1,nrow(mutation.data))
  x1=mutation.data[,15]; x2=mutation.data[,16]
  fit <- glm(count ~ x1+x2, family=poisson())
}
```


only 15.91\% mutations fall within the reference windows. 



## Session information


<!-- Insert the session information into the document -->
```{r session-info}
```
