---
title: "De novo mutations of ASD"
author: "Shengtong  Han"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

## Samples and reference windows


### Reference window

Example_windows_with_div_score.bed: each row is a 50bp window with a few features, such as mutation rate, coding, promoter, GC content, and divergence score; only cover genes and 10-kb regions within TSSs. 


### Case Samples 



index    | Samples | \# mutations| \# subjects
----|-------|-------|---
1 | Jiang_cases_DNM.bed | 2,091| 32
2 | Michaelson_cases_DNM.bed |581 | 10
3 | Yuen_NM2015_cases_DNM.bed | 9,381| 162
4 | Wu_cases_DNM.bed | 1,915 | 32
5 | Kong_cases_DNM.bed | 4,930 | 78

Table: Five WGS  DNM case sample information



### Control sample 

index    | Samples | \# mutations| \# subjects
----|-------|-------|---
1 | Wong_NC_2016_693_control_SNV.bed | 27,092|  693
2 | A->C | 990 (3.65%) | 
3 | A->G | 3,812 (14.07%) |
4 | A->T | 815 (3.01%) |
5 | C->G | 1,274 (4.70%) |
6 | C->T | 5,314 (19.61%) |
7 | C->A | 1,298 (4.79%) |
Table:WGS  DNM control sample information 


### 100 bp window 

For each mutation type in the above table, fit the Poisson regression on the feature DHS 


mutation type| $\hat{\beta}_0$ (intercept) | $\hat{\beta_1}$ (DHS)
---|-----|-------------|------
 A->C | -10.47 | 0.45
 A->G | -9.08 | 0.33
 A->T | -10.62 | 0.30
 C->G | -10.31 | 0.75
 C->T | -8.87 | 0.70
 C->A | -10.24 | 0.60 

## Session information


<!-- Insert the session information into the document -->
```{r session-info}
```
