---
title: "Fitting Diagnostic-Simons control data"
author: "Shengtong  Han"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE}
library(knitr)
library(kableExtra)
library(RColorBrewer)
library(gplots)
library(ape)
library(tidyverse)
set.seed(123)
```


```{r, echo=F}
loadRData <- function(fileName){
#loads an RData file, and returns it
    load(fileName)
    get(ls()[ls() != "fileName"])
}
```



##  100 bp windows 


index  | mutation type | # total windows |# window $p.val<0.05$ |  # window $p.val=0$
----|-------|-------|---|-----------
1 | A->C | 26,839,568 | 3,899 |   3,870
2 | A->G | 15,067 | 15,067 | 10,806
3 | A->T | 26,839,598 | 3,656 | 3,649

```{r, echo=T, eval=F}
Data: /project2/xinhe/Shengtong/mutation_rate/mutation.rate/Simons_WGS/controls/table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation.ExpObsVar.bed.bed.add.100bp
Columns: chr, start, end, observed rate, fitted rate.
```


## 5k windows 

```{r, echo=F, message=F, warning=F, cache=T}
path="C:\\Shengtong\\Research\\mutation-rate\\ObsvsExp.5k.window\\"
ObsandExp=as_tibble(read.table(paste(path, "table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation.A_to_C.and.mutation.T_to_G.ExpObsVar.bed.sorted.bed.added.5k", sep="")))
#ggplot(ObsandExp, aes(x=V4, y=V5)) +
 # geom_point()
plot(as.numeric(as.character(ObsandExp$V5)), as.numeric(as.character(ObsandExp$V4)), lty=2, pch=16, xlab="exp", ylab="obs", ylim=c(0,4), main="A_to_C")
#dat=data.frame(obs=as.numeric(as.character(ObsandExp$V4)), exp=as.numeric(as.character(ObsandExp$V5)))
#dat[is.na(dat)] <- 0 # replace NA with 0 
 #dat=dat[complete.cases(dat), ] # delete rows with NA
# cor(dat[,1], dat[,2])

#dat=data.frame(obs=ObsandExp$V4, exp=ObsandExp$V5)
#cor(as.numeric(dat$obs), as.numeric(dat$exp))
```

* aggregate summary statistics, expected rate and observed count in 100 bp windows 

* there are 2 windows with expected rate >100 

```
> ObsandExp[which(as.numeric(as.character(ObsandExp$V5))>100),]
# A tibble: 2 x 5
  V1           V2        V3 V4    V5         
  <chr>     <int>     <int> <chr> <chr>      
1 chr8  139250000 139255000 0     166.3652155
2 chr8  139255000 139260000 0     101.249892
```

See below  the high recombination rate of two 100 bp windows. 

```
chr8     139249900       139250000       0       5.5363439e-08   0       0       1       0.44    .       108.804         0       0         1       0       0       0       0       0
chr8     139250000       139250100       0       3.779853e-08    0       0       1       0.51    -1.174439       108.804         0         0       1       0       0       0       0       0
```

```{r, echo=F, warning=F, message=F}
outlier_index=which(as.numeric(as.character(ObsandExp$V5))>10)
plot(as.numeric(as.character(ObsandExp$V5[-outlier_index])), as.numeric(as.character(ObsandExp$V4[-outlier_index])), lty=2, pch=16, xlab="exp", ylab="obs", ylim=c(0,4), main="A_to_C with 3 outliers removed")
```




```{r, echo=F, message=F, warning=F, cache=T}
path="C:\\Shengtong\\Research\\mutation-rate\\ObsvsExp.5k.window\\"
ObsandExp=as_tibble(read.table(paste(path, "table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation.C_to_G.nonCpG.and.mutation.G_to_C.nonCpG.ExpObsVar.bed.sorted.bed.observed.added.5k", sep="")))
#ggplot(ObsandExp, aes(x=V4, y=V5)) +
 # geom_point()
plot(as.numeric(as.character(ObsandExp$V5)), as.numeric(as.character(ObsandExp$V4)), lty=2, pch=16, xlab="exp", ylab="obs", ylim=c(0,4), main="C_to_G.nonCpG")
dat=data.frame(obs=as.numeric(as.character(ObsandExp$V4)), exp=as.numeric(as.character(ObsandExp$V5)))
dat[is.na(dat)] <- 0 # replace NA with 0 
 #dat=dat[complete.cases(dat), ] # delete rows with NA
 cor(dat[,1], dat[,2])

#dat=data.frame(obs=ObsandExp$V4, exp=ObsandExp$V5)
#cor(as.numeric(dat$obs), as.numeric(dat$exp))
```



## 10k windows 


add 100bp windows: Observed and estimated mutations for 100k windows are obtained by adding up that in 100bp windows. 


```{r, echo=F}
plotQQ.unif <- function(p.obs) {
  obs <- (p.obs)
  theo <- (ppoints(length(obs)))
  qqplot(-log(theo, base=10), -log(obs, base=10), xlab=expression(paste("Theoretical ",-log[10], "(p-values)")), ylab=expression(paste("Observed ", -log[10], "(p-values)")))
  abline(0,1,col='red')
}
```



### Residual distribution and QQ plot of p values 

```{r, echo=F, warning=F, cache=T}
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
path="C:\\Shengtong\\Research\\mutation-rate\\parameter_estimate\\Simons\\10kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
for (i in 1:length(mut.type))
{
#  i=2
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added.10k", sep=""), header=F)

stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))
pval=pnorm(stand.diff, 0, 1, lower.tail=F)
simons_control_10k=cbind(simons_control, stand.diff, pval)
simons_control_10k=simons_control_10k[complete.cases(simons_control_10k), ]
simons_control_10k[simons_control_10k==Inf]=10^{20}
y.min=min(simons_control_10k[,6])-1; y.max=max(simons_control_10k[,6])+1
par(mfrow=c(1,2))

op <- par(mar = c(5,6,4,2) + 0.1)
plot(simons_control_10k[,6], type="l", ylim=c(y.min, y.max), ylab="", xlab="window index", main=mut.type[i])
abline(h=1.75, col="red")
mtext(expression(frac(y[i]-hat(mu)[i], sqrt(hat(mu)[i]))), side=2, line=2.2, cex=1.2)
grid()
par(op)

plotQQ.unif(simons_control_10k[,7])
grid()
}

```



### Windows with high expected mutation rate 


```{r, echo=F, warning=F, cache=T}
cutoff=-5
high.mut.region=c(NA)
high.mut.type=c(NA)
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
#path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\10kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
for (i in 1:length(mut.type))
{
#  i=1
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added.10k", sep=""), header=F)
stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))

if (length(which(stand.diff<(cutoff)))>0)
{
  
  high.mut.region=rbind(high.mut.region, simons_control[which(stand.diff<(cutoff)),]) 
    for (ii  in 1:length(which(stand.diff<(cutoff)))) # multiple windows may be from the same mutation type
       high.mut.type=rbind(high.mut.type, paste(mut.type[i], ii, sep="_"))
  
}

} # end of i 

high.mut.region=high.mut.region[-1,]
rownames(high.mut.region)=high.mut.type[-1]
colnames(high.mut.region)=c("chr", "start", "end", "observed", "expected")
```


```{r, echo=F}
kable(high.mut.region, caption = "windows with standardized residual less than -5")
```


The reason for high expected mutation rate is that there are about 100 windows of 100 bp that have recombination rates as high as 108.  



### hotspot detection and neighborhood correlation 


One way to define hotsopt is look at standardized residuals that are greater than a threshold, say 4 and study the distribution of selected 'hotspot' windows. 


```{r, echo=F, warning=F, cache=T}
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
cutoff=c(2,3,4)
#path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\10kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
pvalue.mut.type=list()
closest.neigh.mut.type=list()
random.closest.neigh.mut.type=list()
for (i in 1:length(mut.type))
{
#  i=2
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added.10k", sep=""), header=F)
stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))
count=matrix(nrow=4, ncol=length(cutoff))
random.count=matrix(nrow=4, ncol=length(cutoff))
pvalue=matrix(nrow=4, ncol=length(cutoff))
closest.neigh.dist=list()
random.win.closest.neigh.dist=list()
for (ii in 1:length(cutoff))
{
simons_control_hotspot=simons_control[which(stand.diff>cutoff[ii]),]
############## hotspot for Xin's grant 
if (ii==3)
write.csv(simons_control_hotspot, file=paste("C:\\Shengtong\\Research\\mutation-rate\\ForXinGrant\\", "10k.window", "stand.diff.cutoff",cutoff[ii], mut.type[i],"CSV", sep="."))  
##############
wind.index=as.numeric(rownames(simons_control_hotspot))
###############################
wind.result=rle(diff(wind.index))
wind.iso=sum(wind.result$values>1)+1
wind.2neighbor=sum(wind.result$lengths==2 & wind.result$values==1)
wind.3neighbor=sum(wind.result$lengths==3 & wind.result$values==1)
wind.3moreneighbor=sum(wind.result$lengths>3 & wind.result$values==1)
count[,ii]=c(wind.iso, wind.2neighbor, wind.3neighbor, wind.3moreneighbor)/sum(c(wind.iso, wind.2neighbor, wind.3neighbor, wind.3moreneighbor))
################### use simulation of random samples to get expected rate of each case 
  total.num=nrow(simons_control)
  hotspot.num=nrow(simons_control[which(stand.diff>cutoff[ii]),])
  ratio.dist=rep(0, 4)
  num.rep=100
  for (jj in 1:num.rep)
  {
  random.index=sort(sample(seq(1,total.num), hotspot.num))
  random.result=rle(diff(random.index))
  if (length(random.result$lengths)>1)
   { 
    random.iso=sum(random.result$values>1)
    random.2neighbor=sum(random.result$lengths==2 & random.result$values==1)
    random.3neighbor=sum(random.result$lengths==3 & random.result$values==1)
    random.3moreneighbor=sum(random.result$lengths>3 & random.result$values==1)
    ratio.dist=ratio.dist+c(random.iso, random.2neighbor, random.3neighbor, random.3moreneighbor)/(sum(c(random.iso, random.2neighbor, random.3neighbor, random.3moreneighbor))-1)
  }
  if (length(random.result$lengths)==1)
   {
    if (random.result$values>1)
       ratio.dist=ratio.dist+c(1,0,0,0) # only isolated windows 
    if (random.result$values==1 & random.result$lengths==1)
       ratio.dist=ratio.dist+c(0,1,0,0) # only 2neighbor windows
    if (random.result$values==1 & random.result$lengths==2)
       ratio.dist=ratio.dist+c(0,0,1,0) # only 3neighbor windows
    if (random.result$values==1 & random.result$lengths>2)
       ratio.dist=ratio.dist+c(0,0,0,1) # only 3moreneighbor windows
   }
 } # end of jj
  random.count[,ii]=ratio.dist/num.rep
  for (jj in 1:nrow(count))
    pvalue[jj,ii]=poisson.test(round(c(count[jj,ii], random.count[jj,ii])*total.num),r=1, alternative = "greater")$p.value
 ###############################  get null distribution of distance of one hotpspot to its closeest neighbor 
  total.num=nrow(simons_control)
  hotspot.num=nrow(simons_control[which(stand.diff>cutoff[ii]),])
  random.window.index=sort(sample(seq(1,total.num), hotspot.num))
  closest.neigh.dist.indi=numeric()
  for (jj in 1:(length(random.window.index)-1))
      closest.neigh.dist.indi[jj]=random.window.index[jj+1]-random.window.index[jj]
  random.win.closest.neigh.dist[[ii]]=closest.neigh.dist.indi  
 ############################## for each window, calculate the distance to the closest neighbor 
  closest.dist=numeric()
  for (jj in 1:length(wind.index))
  {
   if (jj==1) 
    closest.dist[jj]=wind.index[jj+1]-wind.index[jj]
   if (jj>1 & jj<length(wind.index))
     closest.dist[jj]=min(wind.index[jj+1]-wind.index[jj], wind.index[jj]-wind.index[jj-1])
   if (jj==length(wind.index))
     closest.dist[jj]=wind.index[jj]-wind.index[jj-1]
  }
  closest.neigh.dist[[ii]]=closest.dist
} # end of ii
############
closest.neigh.mut.type[[i]]=closest.neigh.dist
random.closest.neigh.mut.type[[i]]=random.win.closest.neigh.dist
###############
rownames(pvalue)=c(0,1,2, ">2")
colnames(pvalue)=cutoff
#kable(pvalue, caption=mut.type[i])
pvalue.mut.type[[i]]=pvalue

rownames(count)=c(0,1,2, ">2")
colnames(count)=cutoff
log.pvalue=-log(pvalue)
log.pvalue[log.pvalue>5]=5
barcenters=barplot(count, col=c("red", "blue", "green", "yellow"), beside=T, legend=c(0,1,2, ">2"), ylim=c(0,6), xlab="Standardized residual cutoff", ylab="Freq", main=mut.type[i],  args.legend=list(title="No.conse win"))
points(as.vector(barcenters), as.vector(log.pvalue), col=rep(c("red", "blue", "green", "yellow"), 4), pch=16)
} # end of i
```



* No.consecutive windows: 0 means 'isolated' windows , 1 means two-consecutive windows, and etc.  
* with cutoff of 4, all "hotsopt" are isolated and consecutive hotspots are clearly observed until cutoff reduces to 2. 


```{r, echo=F, warning=F}
for (j in 1:length(mut.type))
{  

par(mfrow=c(1,3),oma = c(0, 0, 2, 0))
for (i in 1:3)
{
ave=mean(closest.neigh.mut.type[[j]][[i]])
p.value=ks.test(closest.neigh.mut.type[[j]][[i]], random.closest.neigh.mut.type[[j]][[i]])$p.value
hist(closest.neigh.mut.type[[j]][[i]], freq=F, breaks=50, main=paste(" mean of ", round(ave), ";p.val=", round(p.value, digits=6),sep=""), xlab=cutoff[i])
lines(density(closest.neigh.mut.type[[j]][[i]]), col="red")
lines(density(random.closest.neigh.mut.type[[j]][[i]]), col="green")
abline(v=1, col="blue")

}
  mtext(mut.type[j], outer = TRUE, cex = 1.5)
}  
```


* for every hotspot window with a residual cutoff, look at the distance to its nearest neighbor and plot the histgram of distance and density 
* the mean of the distance is shown in each graph and the blue vertical line is reference line with distance of 1. 
* red line are density of sample distribution and green line is pdf of null distribution of random windows 
* p.val is calculated by ks.test for testing if two distributions are identical or not






###  scatter plot 


```{r, echo=F, warning=F, cache=T}
par(mfrow=c(3,3))
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
chr=paste("chr", seq(1,22), sep="")
#path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\10kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."


for (i in 1:length(mut.type))
{
#  i=1
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added.10k", sep=""), header=F)
chrom.scatter=matrix(nrow=length(chr), ncol=2)
for (ii in 1:length(chr))
{
#  ii=1
##########
scatter.matrix=cbind(as.numeric(as.character(simons_control[simons_control[,1]==chr[ii],4])), as.numeric(as.character(simons_control[simons_control[,1]==chr[ii],5])))
scatter=scatter.matrix[complete.cases(scatter.matrix),]
chrom.scatter[ii,]=colSums(scatter)
}
plot(as.numeric(as.character(chrom.scatter[,2])), as.numeric(as.character(chrom.scatter[,1])), main=mut.type[i], xlab="Expected", ylab="Observed", type="p", pch=16)
abline(0,1, col="red")
grid()
#########
 # end of ii

} # end of i 

```


* Every point is one chromosome where expected mutation rate and observed mutation count in each  10k window are added up. 







## 100k windows 

add 100bp windows: Observed and estimated mutations for 100k windows are obtained by adding up that in 100bp windows. 


#### Residual distribution and QQ plot of p values 

```{r, echo=F, warning=F, cache=T}
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
path="C:\\Shengtong\\Research\\mutation-rate\\parameter_estimate\\Simons\\100kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
for (i in 1:length(mut.type))
{
#  i=2
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added", sep=""), header=F)

stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))
pval=pnorm(stand.diff, 0, 1, lower.tail=F)
simons_control_100k=cbind(simons_control, stand.diff, pval)
simons_control_100k=simons_control_100k[complete.cases(simons_control_100k), ]
simons_control_100k[simons_control_100k==Inf]=10^{20}
y.min=min(simons_control_100k[,6])-1; y.max=max(simons_control_100k[,6])+1
par(mfrow=c(1,2))

op <- par(mar = c(5,6,4,2) + 0.1)
plot(simons_control_100k[,6], type="l", ylim=c(y.min, y.max), ylab="", xlab="window index", main=mut.type[i])
abline(h=1.75, col="red")
mtext(expression(frac(y[i]-hat(mu)[i], sqrt(hat(mu)[i]))), side=2, line=2.2, cex=1.2)
grid()
par(op)

plotQQ.unif(simons_control_100k[,7])
grid()
}

```


### Windows with high expected mutation rate 


```{r, echo=F, warning=F, cache=T}
cutoff=-5
high.mut.region=c(NA)
high.mut.type=c(NA)
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
#path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
for (i in 1:length(mut.type))
{
#  i=1
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added", sep=""), header=F)
stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))
high.mut.region=rbind(high.mut.region, simons_control[which(stand.diff<(cutoff)),])
if (length(which(stand.diff<(cutoff)))>0)
  high.mut.type=rbind(high.mut.type, mut.type[i])
}

high.mut.region=high.mut.region[-1,]
rownames(high.mut.region)=high.mut.type[-1]
colnames(high.mut.region)=c("chr", "start", "end", "observed", "expected")
```


```{r, echo=F}
kable(high.mut.region, caption = "windows with standardized residual less than -5")
```


```{r, echo=F, eval=F}
high.mut.chr8=read.table("D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation.A_to_C.ExpObsVar.bed.139200000-139300000")
par(mfrow=c(1,2),oma = c(0, 0, 2, 0))

plot(high.mut.chr8[,5], high.mut.chr8[,4], xlab="Expected", ylab="Observed", type="p", pch=16, main="A_to_C")
abline(0,1, col="red")
grid()

op <- par(mar = c(5,6,4,2) + 0.1)
plot(high.mut.chr8[,6], type="l", xlab="100 bp windows", main="A_to_C", ylab="")
mtext(expression(frac(y[i]-hat(mu)[i], sqrt(hat(mu)[i]))), side=2, line=2.2, cex=1.2)
grid()
par(op)
mtext("100 bp windows for chr8:139200000-139300000", outer = TRUE, cex = 1.5)
```



### hotspot detection and neighborhood correlation 


One way to define hotsopt is look at standardized residuals that are greater than a threshold, say 4 and study the distribution of selected 'hotspot' windows. 


```{r, echo=F, warning=F, cache=F}
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
cutoff=c(2,3,4)
#path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
pvalue.mut.type=list()
closest.neigh.mut.type=list()
random.closest.neigh.mut.type=list()
ExpObsVar.sum=matrix(0, nrow=31063, ncol=2)
for (i in 1:length(mut.type))
{
#  i=1
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added", sep=""), header=F)
ExpObsVar.sum[,1]=ExpObsVar.sum[,1]+as.numeric(as.character(simons_control[,4])) # sum up observed and expected mutation counts across mutation types 
ExpObsVar.sum[,2]=ExpObsVar.sum[,2]+as.numeric(as.character(simons_control[,5])) # sum up observed and expected mutation counts across mutation types

stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))
count=matrix(nrow=4, ncol=length(cutoff))
random.count=matrix(nrow=4, ncol=length(cutoff))
pvalue=matrix(nrow=4, ncol=length(cutoff))
closest.neigh.dist=list()
random.win.closest.neigh.dist=list()
for (ii in 1:length(cutoff))
{
simons_control_hotspot=simons_control[which(stand.diff>cutoff[ii]),]
############hotspot for Xin's grant 
#if (ii==3)
{  
# save the full list  
#write.csv(simons_control_hotspot, file=paste("C:\\Shengtong\\Research\\mutation-rate\\ForXinGrant\\", "100k.window", "stand.diff.cutoff",cutoff[ii], mut.type[i],"CSV", sep="."))  
  
#simons_control_hotspot[,6]=(as.numeric(as.character(simons_control_hotspot[,4]))-as.numeric(as.character(simons_control_hotspot[,5])))/sqrt(as.numeric(as.character(simons_control_hotspot[,5])))
#simons_control_hotspot_sort=simons_control_hotspot[order(simons_control_hotspot$V6, decreasing=T),]
#write.csv(simons_control_hotspot_sort[1:5,], file=paste("C:\\Shengtong\\Research\\mutation-rate\\ForXinGrant\\", "100k.window", "stand.diff.cutoff",cutoff[ii], "top5window", mut.type[i],"CSV", sep=".")) 
}  
############
wind.index=as.numeric(rownames(simons_control_hotspot))
###############################
wind.result=rle(diff(wind.index))
wind.iso=sum(wind.result$values>1)+1
wind.2neighbor=sum(wind.result$lengths==2 & wind.result$values==1)
wind.3neighbor=sum(wind.result$lengths==3 & wind.result$values==1)
wind.3moreneighbor=sum(wind.result$lengths>3 & wind.result$values==1)
count[,ii]=c(wind.iso, wind.2neighbor, wind.3neighbor, wind.3moreneighbor)/sum(c(wind.iso, wind.2neighbor, wind.3neighbor, wind.3moreneighbor))
################### use simulation of random samples to get expected rate of each case 
  total.num=nrow(simons_control)
  hotspot.num=nrow(simons_control[which(stand.diff>cutoff[ii]),])
  ratio.dist=rep(0, 4)
  num.rep=100
  for (jj in 1:num.rep)
  {
  random.index=sort(sample(seq(1,total.num), hotspot.num))
  random.result=rle(diff(random.index))
  if (length(random.result$lengths)>1)
   { 
    random.iso=sum(random.result$values>1)
    random.2neighbor=sum(random.result$lengths==2 & random.result$values==1)
    random.3neighbor=sum(random.result$lengths==3 & random.result$values==1)
    random.3moreneighbor=sum(random.result$lengths>3 & random.result$values==1)
    ratio.dist=ratio.dist+c(random.iso, random.2neighbor, random.3neighbor, random.3moreneighbor)/(sum(c(random.iso, random.2neighbor, random.3neighbor, random.3moreneighbor))-1)
  }
  if (length(random.result$lengths)==1)
   {
    if (random.result$values>1)
       ratio.dist=ratio.dist+c(1,0,0,0) # only isolated windows 
    if (random.result$values==1 & random.result$lengths==1)
       ratio.dist=ratio.dist+c(0,1,0,0) # only 2neighbor windows
    if (random.result$values==1 & random.result$lengths==2)
       ratio.dist=ratio.dist+c(0,0,1,0) # only 3neighbor windows
    if (random.result$values==1 & random.result$lengths>2)
       ratio.dist=ratio.dist+c(0,0,0,1) # only 3moreneighbor windows
   }
 } # end of jj
  random.count[,ii]=ratio.dist/num.rep
  for (jj in 1:nrow(count))
    pvalue[jj,ii]=poisson.test(round(c(count[jj,ii], random.count[jj,ii])*total.num),r=1, alternative = "greater")$p.value
 ###############################  get null distribution of distance of one hotpspot to its closeest neighbor 
  total.num=nrow(simons_control)
  hotspot.num=nrow(simons_control[which(stand.diff>cutoff[ii]),])
  random.window.index=sort(sample(seq(1,total.num), hotspot.num))
  closest.neigh.dist.indi=numeric()
  for (jj in 1:(length(random.window.index)-1))
      closest.neigh.dist.indi[jj]=random.window.index[jj+1]-random.window.index[jj]
  random.win.closest.neigh.dist[[ii]]=closest.neigh.dist.indi  
 ############################## for each window, calculate the distance to the closest neighbor 
  closest.dist=numeric()
  for (jj in 1:length(wind.index))
  {
   if (jj==1) 
    closest.dist[jj]=wind.index[jj+1]-wind.index[jj]
   if (jj>1 & jj<length(wind.index))
     closest.dist[jj]=min(wind.index[jj+1]-wind.index[jj], wind.index[jj]-wind.index[jj-1])
   if (jj==length(wind.index))
     closest.dist[jj]=wind.index[jj]-wind.index[jj-1]
  }
  closest.neigh.dist[[ii]]=closest.dist
} # end of ii
##############

############
closest.neigh.mut.type[[i]]=closest.neigh.dist
random.closest.neigh.mut.type[[i]]=random.win.closest.neigh.dist
###############
rownames(pvalue)=c(0,1,2, ">2")
colnames(pvalue)=cutoff
#kable(pvalue, caption=mut.type[i])
pvalue.mut.type[[i]]=pvalue

rownames(count)=c(0,1,2, ">2")
colnames(count)=cutoff
log.pvalue=-log(pvalue)
log.pvalue[log.pvalue>5]=5
barcenters=barplot(count, col=c("red", "blue", "green", "yellow"), beside=T, legend=c(0,1,2, ">2"), ylim=c(0,6), xlab="Standardized residual cutoff", ylab="Freq", main=mut.type[i],  args.legend=list(title="No.conse win"))
points(as.vector(barcenters), as.vector(log.pvalue), col=rep(c("red", "blue", "green", "yellow"), 4), pch=16)
} # end of i

################### consider total mutation count acoss all mutation types 
stand.diff=(ExpObsVar.sum[,1]-ExpObsVar.sum[,2])/sqrt(ExpObsVar.sum[,2])
simons_control_hotspot=simons_control[which(stand.diff>4),]
simons_control_hotspot[,6]=stand.diff[which(stand.diff>4)]
simons_control_hotspot[,4]=ExpObsVar.sum[which(stand.diff>4),1]
simons_control_hotspot[,5]=ExpObsVar.sum[which(stand.diff>4),2]
simons_control_hotspot_sort=simons_control_hotspot[order(simons_control_hotspot$V6, decreasing=T),]
write.csv(simons_control_hotspot_sort, file=paste("C:\\Shengtong\\Research\\mutation-rate\\ForXinGrant\\", "100k.window", "stand.diff.cutoff.4", "topwindow","across.all.mutation.type","CSV", sep=".")) 
##################

```


* No.consecutive windows: 0 means 'isolated' windows , 1 means two-consecutive windows, and etc.  
* with cutoff of 4, all "hotsopt" are isolated and consecutive hotspots are clearly observed until cutoff reduces to 2. 
* points over the bars are -log(p values) of enrichment with background, -log(p values) are forced to be 5 when they are greater than 5. 


```{r, echo=F, eval=F}
library(depmixS4)
total.state=rep(0,nrow(simons_control))
total.state[wind.index]=1
total.state=data.frame(total.state)
colnames(total.state)=c("state")
mod=depmix(list(state~1),data=total.state,nstates=2, family = list(multinomial("identity")))
fm=fit(mod)

```



```{r, echo=F, results='asis', warning=F, eval=F}
  kable(pvalue.mut.type[[1]], caption=mut.type[1], "html")%>%
kable_styling() %>%
scroll_box(height = "200px")
  kable(pvalue.mut.type[[2]], caption=mut.type[2], "html")%>%
kable_styling() %>%
scroll_box(height = "200px")
  kable(pvalue.mut.type[[3]], caption=mut.type[3], "html")%>%
kable_styling() %>%
scroll_box(height = "200px")
  kable(pvalue.mut.type[[4]], caption=mut.type[4], "html")%>%
kable_styling() %>%
scroll_box(height = "200px")
  kable(pvalue.mut.type[[5]], caption=mut.type[5], "html")%>%
kable_styling() %>%
scroll_box(height = "200px")
  kable(pvalue.mut.type[[6]], caption=mut.type[6], "html")%>%
kable_styling() %>%
scroll_box(height = "200px")
  kable(pvalue.mut.type[[7]], caption=mut.type[7], "html")%>%
kable_styling() %>%
scroll_box(height = "200px")
  kable(pvalue.mut.type[[8]], caption=mut.type[8], "html")%>%
kable_styling() %>%
scroll_box(height = "200px")
  kable(pvalue.mut.type[[9]], caption=mut.type[9], "html")%>%
kable_styling() %>%
scroll_box(height = "200px")
  
```

* In every table, each column is the residual cutoff and each row is the number of consecutive windows.  
* Numbers in the table are p values: In the whole genome, randomly pick the same amount of random windows as hotspot windows, do a enrichment test between hotspot windows and random windows under each scenario by one-side poisson test. 
  



```{r, echo=F, warning=F}
for (j in 1:length(mut.type))
{  

par(mfrow=c(1,3),oma = c(0, 0, 2, 0))
for (i in 1:3)
{
ave=mean(closest.neigh.mut.type[[j]][[i]])
p.value=ks.test(closest.neigh.mut.type[[j]][[i]], random.closest.neigh.mut.type[[j]][[i]])$p.value
hist(closest.neigh.mut.type[[j]][[i]], freq=F, breaks=50, main=paste(" mean of ", round(ave), ";p.val=", round(p.value, digits=6),sep=""), xlab=cutoff[i])
lines(density(closest.neigh.mut.type[[j]][[i]]), col="red")
lines(density(random.closest.neigh.mut.type[[j]][[i]]), col="green")
abline(v=1, col="blue")

}
  mtext(mut.type[j], outer = TRUE, cex = 1.5)
}  
```


* for every hotspot window with a residual cutoff, look at the distance to its nearest neighbor and plot the histgram of distance and density 
* the mean of the distance is shown in each graph and the blue vertical line is reference line with distance of 1. 
* red line are density of sample distribution and green line is pdf of null distribution of random windows 
* p.val is calculated by ks.test for testing if two distributions are identical or not


###  scatter plot 


```{r, echo=F, warning=F, cache=T}
par(mfrow=c(3,3))
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
chr=paste("chr", seq(1,22), sep="")
#path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."


for (i in 1:length(mut.type))
{
#  i=1
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added", sep=""), header=F)
chrom.scatter=matrix(nrow=length(chr), ncol=2)
for (ii in 1:length(chr))
{
#  ii=1
##########
scatter.matrix=cbind(as.numeric(as.character(simons_control[simons_control[,1]==chr[ii],4])), as.numeric(as.character(simons_control[simons_control[,1]==chr[ii],5])))
scatter=scatter.matrix[complete.cases(scatter.matrix),]
chrom.scatter[ii,]=colSums(scatter)
}
plot(as.numeric(as.character(chrom.scatter[,2])), as.numeric(as.character(chrom.scatter[,1])), main=mut.type[i], xlab="Expected", ylab="Observed", type="p", pch=16)
abline(0,1, col="red")
grid()
#########
 # end of ii

} # end of i 

```


* Every point is one chromosome where expected mutation rate and observed mutation count in each  100k window are added up. 





```{r, echo=F, warning=F, eval=F}
#### Relative risk of each window
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
#path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."


for (i in 1:length(mut.type))
{
#  i=3
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added", sep=""), header=F)

relative.risk=as.numeric(as.character(simons_control[,4]))/as.numeric(as.character(simons_control[,5]))
simons_control_RR=cbind(simons_control, relative.risk)
relative.risk[relative.risk==Inf]=10^20

y.min=min(relative.risk, na.rm=T); y.max=max(relative.risk, na.rm=T)

op <- par(mar = c(5,6,4,2) + 0.1)
cutoff=4
win.ratio=sum(relative.risk>cutoff, na.rm=T)/length(relative.risk)
plot(simons_control_RR[,6], type="p", pch=16, ylim=c(y.min, y.max), ylab="", xlab="window index", main=mut.type[i])
abline(h=cutoff, col="red")
mtext(expression(frac(y[i], hat(mu)[i])), side=2, line=2.2, cex=1.2)
grid()
par(op)

}



```








```{r, echo=F, warning=F, eval=F}
# Moran's I
mut.type=c("A_to_C", "A_to_G", "A_to_T")
#path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
chrom=paste("chr", seq(1,22), sep="")
all.pval=matrix(nrow=length(mut.type), ncol=length(chrom))
for (i in 1:length(mut.type))
{ 
#  i=1
#cat(i, "is running", "\n")
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added", sep=""), header=F)
stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))
simons_control_100k=cbind(simons_control, stand.diff)

simons_control_100k=simons_control_100k[complete.cases(simons_control_100k), ]
auto.pval=numeric()
num.chrom=length(chrom)
for (ii in 1:num.chrom)
{
#  ii=1
  chrom.data=simons_control_100k[simons_control_100k[,1]==chrom[ii],]
  dist=matrix(0, nrow=nrow(chrom.data), ncol=nrow(chrom.data))
  for (aa in 1:(nrow(dist)-1))
    dist[aa,aa+1]=1
  dist.inv=dist+t(dist)
  diag(dist.inv)=0
  moran=Moran.I(chrom.data[,6], dist.inv)
  auto.pval[ii]=moran$p.value
  auto.pval[auto.pval==0]=10^(-20) # set zero pvalue to 10^{-10}
} # end of ii
all.pval[i,]=auto.pval
} # end of i 
max.pval=max(1.5 -log(all.pval, base=10))
plot(seq(1,22), -log(all.pval[1,], base=10), main="", ylim=c(0, max.pval), xlab="chr", xaxt='n', ylab=expression(paste(-log[10](pval)), type="o", lty="solid", col=1))
abline(h=1.3, col="red", lty=3)
axis.labels=seq(1,22)
axis(1, at=seq(1,22), labels=axis.labels, cex.axis=0.8)
for (i in 2:length(mut.type))
 lines(-log(all.pval[i,], base=10), main="", ylim=c(0, max.pval), xlab="", ylab="", type="o", pch=i, lty=1, col=i) 
legend(1, 4, mut.type, col=seq(1, length(mut.type)), lty=rep(1, length(mut.type)), pch=seq(1, length(mut.type)), cex=0.8 )  
grid()



```


















## Session information


<!-- Insert the session information into the document -->
```{r session-info}
```


