---
title: "Fitting Diagnostic"
author: "Shengtong  Han"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE}
knitr::read_chunk("chunks.R")
library(knitr)
library(kableExtra)
library(RColorBrewer)
library(gplots)
library(ape)
set.seed(123)
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r, echo=F}
loadRData <- function(fileName){
#loads an RData file, and returns it
    load(fileName)
    get(ls()[ls() != "fileName"])
}
```



## Simons-control 

###  100 bp windows 


index  | mutation type | # total windows |# window $p.val<0.05$ |  # window $p.val=0$
----|-------|-------|---|-----------
1 | A->C | 26,839,568 | 3,899 |   3,870
2 | A->G | 15,067 | 15,067 | 10,806
3 | A->T | 26,839,598 | 3,656 | 3,649





### 10k windows 


add 100bp windows: Observed and estimated mutations for 100k windows are obtained by adding up that in 100bp windows. 


```{r, echo=F}
plotQQ.unif <- function(p.obs) {
  obs <- (p.obs)
  theo <- (ppoints(length(obs)))
  qqplot(-log(theo, base=10), -log(obs, base=10), xlab=expression(paste("Theoretical ",-log[10], "(p-values)")), ylab=expression(paste("Observed ", -log[10], "(p-values)")))
  abline(0,1,col='red')
}
```



#### Residual distribution and QQ plot of p values 

```{r, echo=F, warning=F, cache=T}
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\10kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
for (i in 1:length(mut.type))
{
#  i=2
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added.10k", sep=""), header=F)

stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))
pval=pnorm(stand.diff, 0, 1, lower.tail=F)
simons_control_10k=cbind(simons_control, stand.diff, pval)
simons_control_10k=simons_control_10k[complete.cases(simons_control_10k), ]
simons_control_10k[simons_control_10k==Inf]=10^{20}
y.min=min(simons_control_10k[,6])-1; y.max=max(simons_control_10k[,6])+1
par(mfrow=c(1,2))

op <- par(mar = c(5,6,4,2) + 0.1)
plot(simons_control_10k[,6], type="l", ylim=c(y.min, y.max), ylab="", xlab="window index", main=mut.type[i])
abline(h=1.75, col="red")
mtext(expression(frac(y[i]-hat(mu)[i], sqrt(hat(mu)[i]))), side=2, line=2.2, cex=1.2)
grid()
par(op)

plotQQ.unif(simons_control_10k[,7])
grid()
}

```



#### Windows with high expected mutation rate 


```{r, echo=F, warning=F, cache=T}
cutoff=-5
high.mut.region=c(NA)
high.mut.type=c(NA)
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\10kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
for (i in 1:length(mut.type))
{
#  i=1
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added.10k", sep=""), header=F)
stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))

if (length(which(stand.diff<(cutoff)))>0)
{
  
  high.mut.region=rbind(high.mut.region, simons_control[which(stand.diff<(cutoff)),]) 
    for (ii  in 1:length(which(stand.diff<(cutoff)))) # multiple windows may be from the same mutation type
       high.mut.type=rbind(high.mut.type, paste(mut.type[i], ii, sep="_"))
  
}

} # end of i 

high.mut.region=high.mut.region[-1,]
rownames(high.mut.region)=high.mut.type[-1]
colnames(high.mut.region)=c("chr", "start", "end", "observed", "expected")
```


```{r, echo=F}
kable(high.mut.region, caption = "windows with standardized residual less than -5")
```


The reason for high expected mutation rate is that there are about 100 windows of 100 bp that have recombination rates as high as 108.  



#### hotspot detection and neighborhood correlation 


One way to define hotsopt is look at standardized residuals that are greater than a threshold, say 4 and study the distribution of selected 'hotspot' windows. 


```{r, echo=F, warning=F, cache=T}
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
cutoff=c(2,3,4)
path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\10kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
for (i in 1:length(mut.type))
{
#  i=2
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added.10k", sep=""), header=F)

stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))
count=matrix(nrow=4, ncol=length(cutoff))
for (ii in 1:length(cutoff))
{
simons_control_hotspot=simons_control[which(stand.diff>cutoff[ii]),]
wind.index=as.numeric(rownames(simons_control_hotspot))
wind.result=rle(diff(wind.index))
wind.iso=sum(wind.result$values>1)+1
wind.2neighbor=sum(wind.result$lengths==2 & wind.result$values==1)
wind.3neighbor=sum(wind.result$lengths==3 & wind.result$values==1)
wind.3moreneighbor=sum(wind.result$lengths>3 & wind.result$values==1)
count[,ii]=c(wind.iso, wind.2neighbor, wind.3neighbor, wind.3moreneighbor)/sum(c(wind.iso, wind.2neighbor, wind.3neighbor, wind.3moreneighbor))
}

rownames(count)=c(0,1,2, ">2")
colnames(count)=cutoff
barplot(count, col=c("red", "blue", "green", "yellow"), beside=T, legend=c(0,1,2, ">2"), ylim=c(0,2), xlab="Standardized residual cutoff", ylab="Freq", main=mut.type[i],  args.legend=list(title="No.consective windows"))
}
```



* No.consecutive windows: 0 means 'isolated' windows , 1 means two-consecutive windows, and etc.  
* with cutoff of 4, all "hotsopt" are isolated and consecutive hotspots are clearly observed until cutoff reduces to 2. 



####  scatter plot 


```{r, echo=F, warning=F, cache=T}
par(mfrow=c(3,3))
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
chr=paste("chr", seq(1,22), sep="")
path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\10kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."


for (i in 1:length(mut.type))
{
#  i=1
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added.10k", sep=""), header=F)
chrom.scatter=matrix(nrow=length(chr), ncol=2)
for (ii in 1:length(chr))
{
#  ii=1
##########
scatter.matrix=cbind(as.numeric(as.character(simons_control[simons_control[,1]==chr[ii],4])), as.numeric(as.character(simons_control[simons_control[,1]==chr[ii],5])))
scatter=scatter.matrix[complete.cases(scatter.matrix),]
chrom.scatter[ii,]=colSums(scatter)
}
plot(as.numeric(as.character(chrom.scatter[,2])), as.numeric(as.character(chrom.scatter[,1])), main=mut.type[i], xlab="Expected", ylab="Observed", type="p", pch=16)
abline(0,1, col="red")
grid()
#########
 # end of ii

} # end of i 

```


* Every point is one chromosome where expected mutation rate and observed mutation count in each  10k window are added up. 







### 100k windows 

add 100bp windows: Observed and estimated mutations for 100k windows are obtained by adding up that in 100bp windows. 


#### Residual distribution and QQ plot of p values 

```{r, echo=F, warning=F, cache=T}
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
for (i in 1:length(mut.type))
{
#  i=2
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added", sep=""), header=F)

stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))
pval=pnorm(stand.diff, 0, 1, lower.tail=F)
simons_control_100k=cbind(simons_control, stand.diff, pval)
simons_control_100k=simons_control_100k[complete.cases(simons_control_100k), ]
simons_control_100k[simons_control_100k==Inf]=10^{20}
y.min=min(simons_control_100k[,6])-1; y.max=max(simons_control_100k[,6])+1
par(mfrow=c(1,2))

op <- par(mar = c(5,6,4,2) + 0.1)
plot(simons_control_100k[,6], type="l", ylim=c(y.min, y.max), ylab="", xlab="window index", main=mut.type[i])
abline(h=1.75, col="red")
mtext(expression(frac(y[i]-hat(mu)[i], sqrt(hat(mu)[i]))), side=2, line=2.2, cex=1.2)
grid()
par(op)

plotQQ.unif(simons_control_100k[,7])
grid()
}

```


#### Windows with high expected mutation rate 


```{r, echo=F, warning=F, cache=T}
cutoff=-5
high.mut.region=c(NA)
high.mut.type=c(NA)
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
for (i in 1:length(mut.type))
{
#  i=1
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added", sep=""), header=F)
stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))
high.mut.region=rbind(high.mut.region, simons_control[which(stand.diff<(cutoff)),])
if (length(which(stand.diff<(cutoff)))>0)
  high.mut.type=rbind(high.mut.type, mut.type[i])
}

high.mut.region=high.mut.region[-1,]
rownames(high.mut.region)=high.mut.type[-1]
colnames(high.mut.region)=c("chr", "start", "end", "observed", "expected")
```


```{r, echo=F}
kable(high.mut.region, caption = "windows with standardized residual less than -5")
```


```{r, echo=F}
high.mut.chr8=read.table("D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation.A_to_C.ExpObsVar.bed.139200000-139300000")
par(mfrow=c(1,2),oma = c(0, 0, 2, 0))

plot(high.mut.chr8[,5], high.mut.chr8[,4], xlab="Expected", ylab="Observed", type="p", pch=16, main="A_to_C")
abline(0,1, col="red")
grid()

op <- par(mar = c(5,6,4,2) + 0.1)
plot(high.mut.chr8[,6], type="l", xlab="100 bp windows", main="A_to_C", ylab="")
mtext(expression(frac(y[i]-hat(mu)[i], sqrt(hat(mu)[i]))), side=2, line=2.2, cex=1.2)
grid()
par(op)
mtext("100 bp windows for chr8:139200000-139300000", outer = TRUE, cex = 1.5)
```



#### hotspot detection and neighborhood correlation 


One way to define hotsopt is look at standardized residuals that are greater than a threshold, say 4 and study the distribution of selected 'hotspot' windows. 


```{r, echo=F, warning=F, cache=T}
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
cutoff=c(2,3,4)
path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
for (i in 1:length(mut.type))
{
#  i=2
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added", sep=""), header=F)

stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))
count=matrix(nrow=4, ncol=length(cutoff))
for (ii in 1:length(cutoff))
{
simons_control_hotspot=simons_control[which(stand.diff>cutoff[ii]),]
wind.index=as.numeric(rownames(simons_control_hotspot))
wind.result=rle(diff(wind.index))
wind.iso=sum(wind.result$values>1)+1
wind.2neighbor=sum(wind.result$lengths==2 & wind.result$values==1)
wind.3neighbor=sum(wind.result$lengths==3 & wind.result$values==1)
wind.3moreneighbor=sum(wind.result$lengths>3 & wind.result$values==1)
count[,ii]=c(wind.iso, wind.2neighbor, wind.3neighbor, wind.3moreneighbor)/sum(c(wind.iso, wind.2neighbor, wind.3neighbor, wind.3moreneighbor))
}

rownames(count)=c(0,1,2, ">2")
colnames(count)=cutoff
barplot(count, col=c("red", "blue", "green", "yellow"), beside=T, legend=c(0,1,2, ">2"), ylim=c(0,2), xlab="Standardized residual cutoff", ylab="Freq", main=mut.type[i],  args.legend=list(title="No.consective windows"))
}
```



* No.consecutive windows: 0 means 'isolated' windows , 1 means two-consecutive windows, and etc.  
* with cutoff of 4, all "hotsopt" are isolated and consecutive hotspots are clearly observed until cutoff reduces to 2. 


####  scatter plot 


```{r, echo=F, warning=F, cache=T}
par(mfrow=c(3,3))
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
chr=paste("chr", seq(1,22), sep="")
path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."


for (i in 1:length(mut.type))
{
#  i=1
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added", sep=""), header=F)
chrom.scatter=matrix(nrow=length(chr), ncol=2)
for (ii in 1:length(chr))
{
#  ii=1
##########
scatter.matrix=cbind(as.numeric(as.character(simons_control[simons_control[,1]==chr[ii],4])), as.numeric(as.character(simons_control[simons_control[,1]==chr[ii],5])))
scatter=scatter.matrix[complete.cases(scatter.matrix),]
chrom.scatter[ii,]=colSums(scatter)
}
plot(as.numeric(as.character(chrom.scatter[,2])), as.numeric(as.character(chrom.scatter[,1])), main=mut.type[i], xlab="Expected", ylab="Observed", type="p", pch=16)
abline(0,1, col="red")
grid()
#########
 # end of ii

} # end of i 

```


* Every point is one chromosome where expected mutation rate and observed mutation count in each  100k window are added up. 





```{r, echo=F, warning=F, eval=F}
#### Relative risk of each window
mut.type=c("A_to_C", "A_to_G", "A_to_T", "C_to_A.CpG", "C_to_A.nonCpG", "C_to_G.CpG", "C_to_G.nonCpG", "C_to_T.CpG", "C_to_T.nonCpG")
path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."


for (i in 1:length(mut.type))
{
#  i=3
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added", sep=""), header=F)

relative.risk=as.numeric(as.character(simons_control[,4]))/as.numeric(as.character(simons_control[,5]))
simons_control_RR=cbind(simons_control, relative.risk)
relative.risk[relative.risk==Inf]=10^20

y.min=min(relative.risk, na.rm=T); y.max=max(relative.risk, na.rm=T)

op <- par(mar = c(5,6,4,2) + 0.1)
cutoff=4
win.ratio=sum(relative.risk>cutoff, na.rm=T)/length(relative.risk)
plot(simons_control_RR[,6], type="p", pch=16, ylim=c(y.min, y.max), ylab="", xlab="window index", main=mut.type[i])
abline(h=cutoff, col="red")
mtext(expression(frac(y[i], hat(mu)[i])), side=2, line=2.2, cex=1.2)
grid()
par(op)

}


# A_to_C: 11.31% windows have $Z>4$
# A_to_G: 2.65%
# A_to_T: 10.89% 
```








```{r, echo=F, warning=F, eval=F}
# Moran's I 
mut.type=c("A_to_C", "A_to_G", "A_to_T")
path="D:\\ResearchWork\\StatisticalGenetics\\Mutation-Rate-project\\parameter_estimate\\Simons\\100kwindow\\"
prefix="table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_mutation."
chrom=paste("chr", seq(1,22), sep="")
all.pval=matrix(nrow=length(mut.type), ncol=length(chrom))
for (i in 1:length(mut.type))
{ 
#  i=1
#cat(i, "is running", "\n")
simons_control=read.table(paste(path, prefix, mut.type[i], ".ExpObsVar.bed.bed.added", sep=""), header=F)
stand.diff=(as.numeric(as.character(simons_control[,4]))-as.numeric(as.character(simons_control[,5])))/sqrt(as.numeric(as.character(simons_control[,5])))
simons_control_100k=cbind(simons_control, stand.diff)

simons_control_100k=simons_control_100k[complete.cases(simons_control_100k), ]
auto.pval=numeric()
num.chrom=length(chrom)
for (ii in 1:num.chrom)
{
#  ii=1
  chrom.data=simons_control_100k[simons_control_100k[,1]==chrom[ii],]
  dist=matrix(0, nrow=nrow(chrom.data), ncol=nrow(chrom.data))
  for (aa in 1:(nrow(dist)-1))
    dist[aa,aa+1]=1
  dist.inv=dist+t(dist)
  diag(dist.inv)=0
  moran=Moran.I(chrom.data[,6], dist.inv)
  auto.pval[ii]=moran$p.value
  auto.pval[auto.pval==0]=10^(-20) # set zero pvalue to 10^{-10}
} # end of ii
all.pval[i,]=auto.pval
} # end of i 
max.pval=max(1.5 -log(all.pval, base=10))
plot(seq(1,22), -log(all.pval[1,], base=10), main="", ylim=c(0, max.pval), xlab="chr", xaxt='n', ylab=expression(paste(-log[10](pval)), type="o", lty="solid", col=1))
abline(h=1.3, col="red", lty=3)
axis.labels=seq(1,22)
axis(1, at=seq(1,22), labels=axis.labels, cex.axis=0.8)
for (i in 2:length(mut.type))
 lines(-log(all.pval[i,], base=10), main="", ylim=c(0, max.pval), xlab="", ylab="", type="o", pch=i, lty=1, col=i) 
legend(1, 4, mut.type, col=seq(1, length(mut.type)), lty=rep(1, length(mut.type)), pch=seq(1, length(mut.type)), cex=0.8 )  
grid()

# on every chromsome, calculate residual auto-spatial correlation by Moran.I function
# use incidence matrix-only neighbor windows has distance of 1, 0 otherwise.   
# significant spatial correlations are observed on chromsome 3, 8, 14, and 20  

```


















## Session information


<!-- Insert the session information into the document -->
```{r session-info}
```


