---
title: "Modeling"
author: "Shengtong Han"
date: YYYY-MM-DD
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## Session information

## initial simple model


At every genomic locus, observed number of mutations, $y_i$ is modeled as Poisson distribution as 

$y_i \sim Pois(\mu_i)$

There are several covariates affecting the mutation rate, such as tri-nucleotide context, parents' age, etc. To relate the mutation rate to covariates $X_i$, Poisson regression could be one way to do that 

$log(E(y_i|X_i))=log(\mu_i)=X_i^T\beta$

Thus 

$\mu_i=e^{X_i^T\beta}$. 



If we already have baseline mutation rate, say $\hat{\mu}_{i0}$, we can adjust 

$\mu_i=\hat{\mu}_{i0} e^{X_i^T\beta}$


Mutation rate at single base may be largely influenced by its local context, thus it would be more reasonable to adjust by region level mutation rate. Consider a genomic window, the prior of its mutation rate is defined as 

$\lambda \sim Ga(\alpha, \alpha)$




<!-- Insert the session information into the document -->
```{r session-info}
```
