---
title: "Simons control WGS"
output: html_document
---


```{r, echo=F, message=F, warning=F}
rm(list=ls())
library(tidyverse)
library(gridExtra)
library(ggpubr)
set.seed(123)
```

## Mutation count partition by mutation types 

WGS Simons control data

index    | mutation type | \# mutations
----|-------|-------|---
1 | A->C (+ T->G) | 3,899+3,900=7,799
2 | A->G (+ T->C) | 15,083+14,919=30,002  
3 | A->T (+ T->A) | 3,664+3,743=7,407
4 | C->G (+ G->C) | 5,393+5,308=10,701
5 | C->T (+ G->A) | 23,977+23,907=47,884 
6 | C->A (+ G->T)| 5,404+5,652=11,056

```{r, echo=F}
slices <- c(7799, 30002, 7404, 10701, 47884, 11056) 
lbls <- c("A->C", "A->G", "A->T", "C->G", "C->T", "C->A")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices,labels = lbls, col=rainbow(length(lbls)),
  	main="Percentage of mutation types in Simons control data ")
```


## Model evalulation 



* Model 0 (M0): 3-mer only  for 9 mutation types. This is implemented by enforcing all parameters to be zero in M1.  

* Model  (M1): 3-mer+features for pooled 9 mutation types 

* Model 2 (M2): 3-mer+features for 9 mutation types

### Predictive density measure 

* Whole human genome (hg19) is evenly split into 100bp windows 
* use even (odd) windows as training data to estimate parameters $\beta$
* use another half odd (even) windows as validation data to make prediction 
* windows with predicted density as `NaN` due to $\widehat{\mu}_j=0$ are removed 

#### M0 vs M2 by mutation types 

```{r, echo=F}
path="C:/Shengtong/Research/mutation-rate/mutation-rate-modeling/data/post.predictive.density"
m0.post.pred=tibble()
m2.post.pred=tibble()
m2.post.pred=as_tibble(read.table(paste(path, "post.pred.for.model.txt", sep="/"), header=T))
m0.post.pred=as_tibble(read.table(paste(path, "post.pred.for.model.with.zero.coefficient.and.intercept.txt", sep="/"), header=T))
m0.m2.combine=bind_rows(m0.post.pred, m2.post.pred)%>%add_column(model=c(rep("M0", 18), rep("M2", 18)))
################## plot them 

ggplot(data=m0.m2.combine%>%filter(train.set=="odd"), aes(x=mutation.type, y=-post.pred, fill=model)) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=5))+
theme(text = element_text(size=10))+
geom_bar(stat="identity", position=position_dodge())+
   ylab("-log(pred.prob)")+
   ggtitle("odd widow as train set and even as validation")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title

ggplot(data=m0.m2.combine%>%filter(train.set=="even"), aes(x=mutation.type, y=-post.pred, fill=model)) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size=5))+
theme(text = element_text(size=10))+
geom_bar(stat="identity", position=position_dodge())+
    ylab("-log(pred.prob)")+
   ggtitle("even widow as train set and odd as validation")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title
```

#### three models regardless of mutation types 

* M0, M2 sum over all predicted probabilities across 9 mutation types  



```{r, echo=F}
m1.post.pred=as_tibble(read.table(paste(path, "pooled.mutation.type.log.predictive.density.txt", sep="/"), header=T))
pp0=m0.post.pred%>%select(post.pred)%>%summarise(sum=sum(post.pred))%>%pull()
pp1=m1.post.pred%>%select(post.pred)%>%summarise(sum=sum(post.pred))%>%pull()
pp2=m2.post.pred%>%select(post.pred)%>%summarise(sum=sum(post.pred))%>%pull()  
total.post.pred=tibble(model=c("M0", "M1", "M2"), post.pred=c(pp0, pp1, pp2))

ggplot(data=total.post.pred, aes(x=model, y=-post.pred)) +
   theme(axis.text.x = element_text(angle = 60, vjust = 0.5, size=10))+
theme(text = element_text(size=10))+
geom_bar(stat="identity", fill="blue")+
    ylab("-log(pred.prob)")+
   #ggtitle("even widow as train set and odd as validation")+
  theme(plot.title = element_text(hjust = 0.5, size=10)) # center the title
```

### Expected rate vs observed rate 


#### M0 vs M2 by mutation types


```{r, echo=F, message=F, warning=F, cache=T}
path="C:/Shengtong/Research/mutation-rate/ObsvsExp.10k.window"
mutation.type=unique(as.character(read.table("C:/Shengtong/Research/mutation-rate/mutation-rate-modeling/data/post.predictive.density/post.pred.for.model.txt", header=T)$mutation.type)) # get mutation types 
Train=c("odd", "even")
figures=list(); fig=0
for (mut in 1:length(mutation.type))
   for (set in 1:length(Train))
   {
     # mut=1; set=1
      m2.exp.obs=as_tibble(read.table(paste(path, "/", "table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_", mutation.type[mut], ".ExpObsVar.", Train[set], ".window.as.train.bed.bed.added.10k", sep=""), header=F))
      
m0.exp.obs=as_tibble(read.table(paste(path, "/", "table.ASCWGS_20180504.WGS1902_hg19_controls_SNV_remove_recurrent_", mutation.type[mut], ".ExpObsVar.", Train[set], ".window.as.train.with.zero.coefficient.and.intercept.bed.bed.added.10k", sep=""), header=F))
colnames(m2.exp.obs)=c("chr", "start", "end", "obs", "exp")
colnames(m0.exp.obs)=c("chr", "start", "end", "obs", "exp")

############### consider chromosome level analysis 
chr=unique(m2.exp.obs$chr)
m0.exp.obs.chr=tibble()
m2.exp.obs.chr=tibble()
for (cc in 1:22) #only look at chr1, chr2, ..., chr22
{
   
   m0.exp.obs.chr[cc,1]=sum(as.numeric(as.character(m0.exp.obs[which(m0.exp.obs$chr==chr[cc]),]$obs)), na.rm=T)
   m0.exp.obs.chr[cc,2]=sum(as.numeric(as.character(m0.exp.obs[which(m0.exp.obs$chr==chr[cc]),]$exp)), na.rm=T)
   
   m2.exp.obs.chr[cc,1]=sum(as.numeric(as.character(m2.exp.obs[which(m2.exp.obs$chr==chr[cc]),]$obs)), na.rm=T)
   m2.exp.obs.chr[cc,2]=sum(as.numeric(as.character(m2.exp.obs[which(m2.exp.obs$chr==chr[cc]),]$exp)), na.rm=T)
   
} 
colnames(m0.exp.obs.chr)=c("obs", "exp")
colnames(m2.exp.obs.chr)=c("obs", "exp")

m0.m2.exp.obs.chr=rbind(m0.exp.obs.chr, m2.exp.obs.chr) %>% add_column(model=c(rep("M0", nrow(m0.exp.obs.chr)), rep("M2", nrow(m2.exp.obs.chr))))  # combine two models together 

x.limit=y.limit=max(m0.m2.exp.obs.chr[,1:2])
fig=fig+1
figures[[fig]]=ggplot(m0.m2.exp.obs.chr, aes(x=obs, y=exp, color=model)) +
  geom_point()+
   xlim(c(0, x.limit))+ylim(c(0, y.limit))+
   geom_abline(intercept = 0, slope = 1)+
   ggtitle(paste(mutation.type[mut]))+
   theme(plot.title = element_text(hjust = 0.5, size=7)) # center the title
}  # end of for (set in 1:length(Train))

figure=ggarrange(figures[[1]],figures[[3]], figures[[5]], figures[[7]], figures[[9]], figures[11], figures[[13]], figures[[15]], figures[[17]],  common.legend = TRUE, legend="right") 
#figure
annotate_figure(figure,
                top = text_grob("odd windows", color = "red", face = "bold", size = 14),
#                bottom = text_grob("Data source: \n ToothGrowth data set", color = "blue",
#                                   hjust = 1, x = 1, face = "italic", size = 10),
#                left = text_grob("Tooth length", color = "green", rot = 90),
#                right = "I'm done, thanks :-)!",
#                fig.lab = "Figure 1", fig.lab.face = "bold"
)


figure=ggarrange(figures[[2]],figures[[4]], figures[[6]], figures[[8]], figures[[10]], figures[12], figures[[14]], figures[[16]], figures[[18]],  common.legend = TRUE, legend="right") 
#figure
annotate_figure(figure,
                top = text_grob("even windows", color = "red", face = "bold", size = 14),
#                bottom = text_grob("Data source: \n ToothGrowth data set", color = "blue",
#                                   hjust = 1, x = 1, face = "italic", size = 10),
#                left = text_grob("Tooth length", color = "green", rot = 90),
#                right = "I'm done, thanks :-)!",
#                fig.lab = "Figure 1", fig.lab.face = "bold"
)
```



## Referecnce 

[Genomic Patterns of De Novo Mutation in Simplex Autism](https://pubmed.ncbi.nlm.nih.gov/28965761/), cell, 2017